---
title: "Problem Section 3"
subtitle:  "Mon Oct 16, 2023"
graphics: yes
output: 
        pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

* * * 
\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence.
The specific tasks include: 

   - Identifying the range of discrete and continuous random variables
   
   - Deriving the PMF and CDF of a discrete random variable
   
   - Recognizing properties of CDFs
   
   - Back up and support work with relevant explanations

\end{shaded}


* * *

### Exercises 


1. For each situation, give the range of the random variable $X$. Then state whether the random variable is discrete or continuous.

a. Three fair coins are tossed independently until they show the same side. $X$ is the number of tosses needed.

The range for this event is $\{1, 2, 3, \dots, \}$ where $X$ is a set of three tosses of the coins. The random variable $X$ is discrete.

b. Using modern aptitude testing and screening procedures, Rote Industries has greatly reduced its employee turnover. Recent records show that 90% of the company’s new assembly line workers stay with the company for at least a year. $X$ is the number of employees among 12 recently hired that will still be with the company in another year.

The range for the random variable $X$ is $\{0, 1, 2, \dots, 12\}$. The random variable is discrete.
		
c. In a busy 24 hour copy center, minor malfunctions of the copy machine occur randomly. $X$ denotes the time (in hours) until the first minor malfunction occurs.

The range for the random variable $X$ is $\{x: x \in \mathbb{R}, x\geq 0\}$ OR we can write $X \in [0, \infty)$. This random variable is continuous.
	 
2. Suppose you toss a fair coin 3 times. The sample space for this experiment is shown below:
$$S = \{ HHH, HHT, HTH, HTT, TTT,TTH, THT, THH \}.$$

a. Write the PMF of $X$ defined as the difference between number of heads and number of tails.

The range of our random variable $X$ is $\{-3, -1, 1, 3\}$ so we can write our PMF as a table with these values of $X$ as the inputs to the function:
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
$x$ & -3 & -1 & 1 & 3 \\
\hline
$f(x)$ & $\frac{1}{8}$ & 3/8 & 3/8 & 1/8 \\
\hline
\end{tabular}
\end{center}

b. Give the CDF of $X$.

\begin{align*}
F(x) &= P(X \leq x) \\
     &= \left\{
     \begin{array}{cc}
        0 &\text{ if } x < -3 \\
        \frac{1}{8} &\text{ if } -3 \leq x < -1 \\
        \frac{1}{2} &\text{ if } -1 \leq x < 1 \\
        \frac{7}{8} &\text{ if } 1 \leq x < 3 \\
        1 &\text{ if } x \geq 3
     \end{array} \right.
\end{align*}

c. Draw the probability histogram of $X$ in R.

```{r ggplot_hist}
pmf_df <- data.frame(
  x = c(-3, -1, 1, 3),
  f = c(1/8, 3/8, 3/8, 1/8)
)

ggplot(data=pmf_df, aes(x=x, y=f)) +
  geom_col(width=1, alpha=0.5,
           fill="gold", color="purple") +
  geom_text(mapping = aes(label=round(f, 4), y = f + 0.01)) +
  labs(title="Probability Histogram of X",
       y = "Probability of X",
       x = "X=x") +
  theme_minimal()
```


	  
3. Which of the following is a valid CDF? (choose one; also point out what's wrong with the ones you didn't choose)

a. 
\begin{align*} 
        F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x \leq 3 \\
                 \frac{3}{4} & 3 <  x < 5 \\
                 1 & 5 \leq x \end{array} \right. \end{align*}
This is not a valid CDF because it is not right-continuous.

b.
\begin{align*} 
        F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x < 3 \\
                 \frac{3}{4} & 3 \leq x < 5 \\
                 1 & 5 \leq x \end{array} \right. \end{align*}
\textbf{This is a valid CDF!}

c. 
\begin{align*} 
      F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{3}{4} & 0 \leq x < 3 \\
                 \frac{1}{4} & 3 \leq x < 5 \\
                 1 & 5 \leq x \end{array} \right. \end{align*}
This is not a valid CDF because it decreases at $x=3$.

d.
\begin{align*} 
       F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x < 3 \\
                 \frac{3}{4} & 3 \leq x < 5 \\
                 \frac{7}{8} & 5 \leq x \end{array}   
                 \right.                      
\end{align*}              
This is not a valid CDF because it does not converge to $1$ as $x \rightarrow \infty$.
